The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-02-14 article/article.tex
conclusion

\subsection{Single-choice balls-into-bins}
\label{sec:single-choice}
A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy. Using this strategy, the maximum load of the fullest bin will be
\begin{comment}
\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.}
\end{comment}
\begin{align*}
\load = \frac{\ln n}{\ln \ln n} \cdot \left(1 + o\left(1\right)\right)
\end{align*}
for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] \geq 1 - n^{-\alpha}$ for an arbitrary constant $\alpha > 0$.}~\cite{RS98}. This implies\footnote{For a detailed proof, see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic}.}
\begin{align*}
\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
\end{align*}
Note that since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is at most~\cite{RS98}.
\begin{align*}
\delta = \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
\end{align*}

@@

\section{Conclusion}
\label{sec:conclusion}

>>>>

\subsection{Single-choice balls-into-bins}
\label{sec:single-choice}
A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy. Using this strategy, the maximum load of the fullest bin will be
\begin{comment}
\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.}
\end{comment}
\begin{align*}
\load = \frac{\ln n}{\ln \ln n} \cdot \left(1 + o\left(1\right)\right)
\end{align*}
for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] \geq 1 - n^{-\alpha}$ for an arbitrary constant $\alpha > 0$.}~\cite{RS98}. This implies\footnote{For a detailed proof, see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic}.}
\begin{align*}
\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
\end{align*}
Note that since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is at most~\cite{RS98}.
\begin{align*}
\delta = \Theta\left(\sqrt{\frac{m \cdot \ln n}{n}}\right).
\end{align*}

@@

\section{Conclusion}
\label{sec:conclusion}
Balls-into-bins games have many applications in Computer Science and have been studied extensively. The single-choice balls-into-bins strategy chooses a bin uniformly and independently at random for each ball. The classical result for this scheme upper-bounds the maximum load with 
\begin{align*}
\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
\end{align*}
w.h.p.
In this paper we have presented two different miltiple-choice balls-into-bins placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
\begin{align*}
\load = \frac{m}{n} + \frac{\ln\ln n}{\ln d} + O(1)
\end{align*}
w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better maximum load of
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{d\cdot \ln \Phi_d} + O(1)
\end{align*}
w.h.p.~for all $m$. We used \emph{witness trees}, a very useful technique in the balls-into-bins context, to prove these upper-bounds in detail.
